Safe and Robust Formal Synthesis of Digital Controllers\\ for Continuous Plants

### 2018-03-17 automatica/paper.tex
edits intro

In this paper, we examine the case of Linear Time Invariant (LTI)
models, for which the synthesis of controllers is well understood.
However, the use of digital control architectures adds new challenges
caused by artifacts specific to digital control, such as the effects
of finite-precision arithmetic, time discretization and quantization
noise introduced by A/D and D/A conversion.
%
Given an LTI model, we develop an automatic technique for generating
correct by construction digital controllers that addresses all these
challenges. Specifically, we synthesize stable and safe,
software-implemented embedded controllers.  Due to the complexity of
such closed-loop systems, we focus on linear models with known
configurations, and perform parametric synthesis of controllers.
%meeting a number of properties that define their desired behavior.

@@

%% Using state of the art verifiers, we evaluate these models against
%% automatically synthesised controllers using CEGIS-based synsthesis that
%% incorporate the above mentioned artifacts in different models of
%% Single-Input/Single-Output (SISO) systems.
In this paper we present two novel approaches for the synthesis of
digital controllers, both of which make use of CounterExample-Guided
Inductive Synthesis (CEGIS)~\cite{jha-icse10,
  DBLP:conf/asplos/Solar-LezamaTBSS06}, a technique
that became very popular in the recent years.  CEGIS denotes an
iterative process, where each iteration performs inductive
generalisation based on counterexamples provided by a verification
oracle. Essentially, the inductive generalisation uses information
about a limited number of inputs to make claims about all the possible
inputs in the form of candidate solutions. Our two instantiations of
CEGIS are described next.

%% offer three modes of
%% operation that address different aspects of the problem.
%% {\em The first  approach} uses a transfer-function based analysis that
%% solves the problem of robust stability for closed loop systems for a
%% collection of unstable plants.  Inthis case, there is no safety
%% specification since we are working in the frequency domain.
{\em The first approach} uses a na\"ive technique that starts by devising a
digital controller that stabilizes the system while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded safety by unfolding the dynamics of the system, considering the
full hyper-cube of initial states, and checking a {\em completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
system such that the boundaries are not violated for any larger number of
iterations.  As~it requires unfolding up to the completeness threshold, this
approach is computationally expensive.

Instead of unfolding the dynamics, {\em the second approach}
employs {\em abstract acceleration}~\cite{cattaruzza2015unbounded} to
evaluate all possible progressions of the system simultaneously. 
Additionally, this approach uses {\em abstraction refinement},
enabling us to always start with a very simple description regardless of the
dynamics complexity, and only expand to more complex models
when a solution cannot be found.

@@

%% \addtodo{CD: the novelty compared to CAV'17 needs to be clarified}
This paper represents an extension of \cite{DBLP:conf/cav/AbateBCCDKKP17}
and makes the following contributions, where the fourth and the fifth denote
extensions over \cite{DBLP:conf/cav/AbateBCCDKKP17}.
%
\begin{enumerate}

  % Additions from CAV:
  % continuous time
  % floating point 
\item We automatically generate {\em correct-by-construction} digital
  controllers using an inductive synthesis approach.  Our application of program
  synthesis is non-trivial and addresses challenges specific to control
  systems, such as quantizers and FWL.  %% In particular, we
  %% have found that a two-stage verification engine that continuously refines
  %% the precision of the fixed-point representation of the plant yields a
  %% speed-up of two orders of magnitude over a conventional one-stage
  %% verification engine.
%
  %\item  We compute
  The computed state-feedback controllers 
  guarantee a given safety property alongside stability and safety
  specifications. Existing methods for controller synthesis rely on transfer
  function representations, which are inadequate to prove safety requirements.
%
\item We provide two novel algorithms: the first, na\"ive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant.
%
\item We develop a model for different sources of quantization error and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.

\item While \cite{DBLP:conf/cav/AbateBCCDKKP17} is restricted to systems
  modeled in discrete time, in the current paper we present a model for
  soundly evaluating reachability in a combined continuous-discrete time/space.
\item A further limitation of the work in \cite{DBLP:conf/cav/AbateBCCDKKP17}
  is its restriction to fixed-point arithmetic, meaning that fixed-point numbers
  were employed for the representation of both the plant and the controller, as well as for
  the operations performed by each of them.
  Conversely, in the current paper we also make use of floating-point arithmetic
  and compare the two approaches. Additionally, we enhance the implementation with an SMT backend, allowing us to also use reals in order to represent the states of the plant.
\end{enumerate}

@@

\textbf{CEGIS -} Program synthesis is the problem of computing
correct-by-design programs from high-level specifications. Algorithms
for this problem have made substantial progress in recent years, in
particular CEGIS, which is a recent approach to inductive synthesis
borrowing ideas from Counter-Example Guided Abstraction Refinement
(CEGAR)~\cite{DBLP:conf/cav/ClarkeGJLV00} and
superoptimisation~\cite{DBLP:conf/iclp/BrainCVF06,
  DBLP:conf/asplos/Massalin87}.  This approach uses an iterative
process where each iteration performs inductive generalisation based
on counterexamples provided by a verification oracle.  In the absence
of complete specifications, oracles are queried to distinguish between
non-equivalent
programs~\cite{jha-icse10,DBLP:conf/fmcad/AlurBJMRSSSTU13}.

>>>>

In this paper, we examine the case of Linear Time Invariant (LTI)
models, for which the synthesis of controllers is well understood.
However, the use of digital control architectures adds new challenges
caused by artifacts specific to digital control, such as the effects
of finite-precision arithmetic, time discretization and quantization
noise introduced by A/D and D/A conversion.
%
Given an LTI model, we develop an automatic technique for generating
correct-by-construction digital controllers that addresses all these
challenges. Specifically, we synthesize stable and safe,
software-implemented embedded controllers.  Due to the complexity of
such closed-loop systems, we focus on linear models with known
configurations, and perform parametric synthesis of controllers.
%meeting a number of properties that define their desired behavior.

@@

%% Using state of the art verifiers, we evaluate these models against
%% automatically synthesised controllers using CEGIS-based synsthesis that
%% incorporate the above mentioned artifacts in different models of
%% Single-Input/Single-Output (SISO) systems.
In this paper we present two novel approaches for the synthesis of
digital controllers, both of which make use of CounterExample-Guided
Inductive Synthesis (CEGIS)~\cite{jha-icse10,
  DBLP:conf/asplos/Solar-LezamaTBSS06}, a technique
that became very popular in the recent years.  CEGIS is an
iterative process, where each iteration performs inductive
generalisation based on counterexamples provided by a verification
oracle. Essentially, the inductive generalisation uses information
about a limited number of inputs to compute a candidate solution
for all the possible. Our two instantiations of CEGIS are
described next.

%% offer three modes of
%% operation that address different aspects of the problem.
%% {\em The first  approach} uses a transfer-function based analysis that
%% solves the problem of robust stability for closed loop systems for a
%% collection of unstable plants.  Inthis case, there is no safety
%% specification since we are working in the frequency domain.

\emph{The first approach} uses a na\"ive technique that starts by devising a
digital controller that stabilizes the system while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded safety by unfolding the dynamics of the system, considering the
full hyper-cube of initial states, and checking a \emph{completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
system such that the boundaries are not violated for any larger number of
iterations.  As~it requires unfolding up to the completeness threshold, this
approach is computationally expensive.

Instead of unfolding the dynamics, \emph{the second approach}
employs \emph{abstract acceleration}~\cite{cattaruzza2015unbounded} to
evaluate all possible progressions of the system simultaneously. 
Additionally, this approach uses \emph{abstraction refinement},
enabling us to always start with a very simple description regardless of the
dynamics complexity, and only expand to more complex models
when a solution cannot be found.

@@

%% \addtodo{CD: the novelty compared to CAV'17 needs to be clarified}
This paper is an extension of~\cite{DBLP:conf/cav/AbateBCCDKKP17}
and makes the following contributions, where the fourth and the fifth are
extensions over~\cite{DBLP:conf/cav/AbateBCCDKKP17}.
%
\begin{enumerate}

  % Additions from CAV:
  % continuous time
  % floating point 
\item We automatically generate \emph{correct-by-construction} digital
  controllers using an inductive synthesis approach.  Our application of program
  synthesis is non-trivial and addresses challenges specific to control
  systems, such as quantizers and FWL.  %% In particular, we
  %% have found that a two-stage verification engine that continuously refines
  %% the precision of the fixed-point representation of the plant yields a
  %% speed-up of two orders of magnitude over a conventional one-stage
  %% verification engine.
%
  %\item  We compute
  The computed state-feedback controllers 
  guarantee a given safety property alongside stability and safety
  specifications. Existing methods for controller synthesis rely on transfer
  function representations, which are inadequate to prove safety requirements.
%
\item We provide two novel algorithms: the first, na\"ive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant.
%
\item We develop a model for different sources of quantization error and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.

\item While~\cite{DBLP:conf/cav/AbateBCCDKKP17} is restricted to systems
  modeled in discrete time, in the current paper we present a model for
  soundly evaluating reachability in a combined continuous-discrete time/space.

\item A further limitation of the work in~\cite{DBLP:conf/cav/AbateBCCDKKP17}
  is its restriction to fixed-point arithmetic, meaning that fixed-point
  numbers were employed for the representation of both the plant and the
  controller, as well as for the operations performed by each of them. 
  Conversely, in the current paper we also make use of floating-point
  arithmetic and compare the two approaches.  Additionally, we enhance the
  implementation with an SMT backend, allowing us to also use reals in order
  to represent the states of the plant.

@@

\textbf{CEGIS --} Program synthesis is the problem of computing
correct-by-design programs from high-level specifications. Algorithms
for this problem have made substantial progress in recent years, in
particular CEGIS, which is a recent approach to inductive synthesis
borrowing ideas from Counter-Example Guided Abstraction Refinement
(CEGAR)~\cite{DBLP:conf/cav/ClarkeGJLV00} and
superoptimisation~\cite{DBLP:conf/iclp/BrainCVF06,
  DBLP:conf/asplos/Massalin87}.  This approach uses an iterative
process where each iteration performs inductive generalisation based
on counterexamples provided by a verification oracle.  In the absence
of complete specifications, oracles are queried to distinguish between
non-equivalent
programs~\cite{jha-icse10,DBLP:conf/fmcad/AlurBJMRSSSTU13}.
